---
title: "Quiz 04"
author: "Matheus Cardoso"
date: "Jun 12, 2020"
output: 
  html_document: 
    fig_caption: yes
    highlight: zenburn
    keep_md: yes
    theme: simplex
    toc: yes
    toc_float: yes
  pdf_document: 
    toc: yes
    fig_caption: yes
    highlight: zenburn
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE, cache = TRUE, warning = FALSE, message = FALSE)

library(caret)
library(tidyverse)
library(rattle)
```

# Q1 

For this quiz we will be using several R packages. R package versions change over time, the right answers have been checked using the following versions of the packages.

AppliedPredictiveModeling: v1.1.6

caret: v6.0.47

ElemStatLearn: v2012.04-0

pgmm: v1.1

rpart: v4.1.8

gbm: v2.1

lubridate: v1.3.3

forecast: v5.6

e1071: v1.6.4

If you aren't using these versions of the packages, your answers may not exactly match the right answer, but hopefully should be close.

Load the vowel.train and vowel.test data sets:

```{r q1}
library(ElemStatLearn)
data(vowel.train)
data(vowel.test)
```

Set the variable y to be a factor variable in both the training and test set. Then set the seed to 33833. Fit (1) a random forest predictor relating the factor variable y to the remaining variables and (2) a boosted predictor using the "gbm" method. Fit these both with the train() command in the caret package.

What are the accuracies for the two approaches on the test data set? What is the accuracy among the test set samples where the two methods agree? 


## A1

```{r a1}
set.seed(33833)

vowel.train[["y"]] <- vowel.train[["y"]] %>% factor()
vowel.test[["y"]]  <- vowel.test[["y"]] %>% factor()

if (!exists("a1_model_rf")){
    a1_model_rf  <- train(y ~ ., method = "rf", data = vowel.train)
}
if (!exists("a1_model_gbm")){
    a1_model_gbm <- train(y ~ ., method = "gbm", data = vowel.train, verbose = F)
}

a1_predict_rf  <- predict(a1_model_rf, vowel.test)
a1_predict_gbm <- predict(a1_model_gbm, vowel.test)

confusionMatrix(a1_predict_rf, vowel.test$y)
confusionMatrix(a1_predict_gbm, vowel.test$y)
confusionMatrix(a1_predict_rf, a1_predict_gbm)
```

# Q2

Load the Alzheimer's data using the following commands

```{r}
library(caret)
library(gbm)
set.seed(3433)
library(AppliedPredictiveModeling)
data(AlzheimerDisease)
adData = data.frame(diagnosis,predictors)
inTrain = createDataPartition(adData$diagnosis, p = 3/4)[[1]]
training = adData[ inTrain,]
testing = adData[-inTrain,]
```

Set the seed to 62433 and predict diagnosis with all the other variables using a random forest ("rf"), boosted trees ("gbm") and linear discriminant analysis ("lda") model. Stack the predictions together using random forests ("rf"). What is the resulting accuracy on the test set? Is it better or worse than each of the individual predictions?

```{r a3}
set.seed(62433)

a2_model_rf <- train(diagnosis ~ ., data = training, method = "rf")
a2_model_gbm <- train(diagnosis ~ ., data = training, method = "gbm", verbose = F)
a2_model_lda <- train(diagnosis ~ ., data = training, method = "lda")

a2_predict_rf <- predict(a2_model_rf, testing)
a2_predict_gbm <- predict(a2_model_gbm, testing)
a2_predict_lda <- predict(a2_model_lda, testing)

a2_combined_df <- data.frame(a2_predict_rf, a2_predict_gbm, a2_predict_lda, diagnosis = testing$diagnosis)

a2_combined_model <- train(diagnosis ~ ., data = a2_combined_df, method = "rf")
a2_combined_predict <- predict(a2_combined_model, a2_combined_df)

confusionMatrix(testing$diagnosis, a2_predict_rf)
confusionMatrix(testing$diagnosis, a2_predict_gbm)
confusionMatrix(testing$diagnosis, a2_predict_lda)
confusionMatrix(testing$diagnosis, a2_combined_predict)
```

## A2

The bias is larger and the variance is smaller. Under leave one out cross validation K is equal to the sample size. 

# Q3

Load the olive oil data using the commands:

```{r q3}
library(pgmm)
data(olive)
olive = olive[,-1]
```

(NOTE: If you have trouble installing the pgmm package, you can download the -code-olive-/code- dataset here: olive_data.zip. After unzipping the archive, you can load the file using the -code-load()-/code- function in R.)

These data contain information on 572 different Italian olive oils from multiple regions in Italy. Fit a classification tree where Area is the outcome variable. Then predict the value of area for the following data frame using the tree command with all defaults

```{r a3}
newdata = as.data.frame(t(colMeans(olive)))

str(newdata)

a2_model <- train(Area ~ ., data = olive, method = "rpart")
predict(a2_model)
```

## A3

2.783. It is strange because Area should be a qualitative variable - but tree is reporting the average value of Area as a numeric variable in the leaf predicted for newdata

# Q4

Load the South Africa Heart Disease Data and create training and test sets with the following code:

```{r q4}
library(ElemStatLearn)
data(SAheart)
set.seed(8484)
train = sample(1:dim(SAheart)[1],size=dim(SAheart)[1]/2,replace=F)
trainSA = SAheart[train,]
testSA = SAheart[-train,]
```

Then set the seed to 13234 and fit a logistic regression model (method="glm", be sure to specify family="binomial") with Coronary Heart Disease (chd) as the outcome and age at onset, current alcohol consumption, obesity levels, cumulative tabacco, type-A behavior, and low density lipoprotein cholesterol as predictors. Calculate the misclassification rate for your model using this function and a prediction on the "response" scale:
```{r}
missClass = function(values,prediction){sum(((prediction > 0.5)*1) != values)/length(values)}
```

```{r a4}
set.seed(13234)
a4_model <- train(
                chd ~ age + alcohol + obesity + tobacco + typea + ldl,
                data = trainSA,
                method = "glm",
                family = "binomial"
)

missClass(trainSA, predict(a4_model, newdata = trainSA))
missClass(testSA, predict(a4_model, newdata = testSA))
```

## A4

This code does not seems to work, even it's apparently right. maybe the data changed because I needed to install the `ElemStatLearn` package from github? (this package isn't available from CRAN anymore...)

# Q5

Load the Alzheimer's disease data using the commands:

```{r q5}
library(caret)
library(AppliedPredictiveModeling)
set.seed(3433)
data(AlzheimerDisease)
adData = data.frame(diagnosis,predictors)
inTrain = createDataPartition(adData$diagnosis, p = 3/4)[[1]]
training = adData[ inTrain,]
testing = adData[-inTrain,]
```

Create a training data set consisting of only the predictors with variable names beginning with IL and the diagnosis. Build two predictive models, one using the predictors as they are and one using PCA with principal components explaining 80% of the variance in the predictors. Use method="glm" in the train function.

What is the accuracy of each method in the test set? Which is more accurate?

```{r a5}
str(training)
names(training)

training_subset <- training %>% 
                        select(starts_with("IL") | "diagnosis")

testing_subset <- testing %>% 
                        select(starts_with("IL") | "diagnosis")

names(training_subset)
names(testing_subset)

# model Non-PCA
mod1 <- train(
            diagnosis ~ .,
            data = training_subset,
            method = "glm"
            )

pred1 <- predict(mod1, testing_subset)

result1 <- confusionMatrix(pred1, testing_subset$diagnosis)

# model PCA
mod2 <- train(
            diagnosis ~ .,
            data = training_subset,
            trControl = trainControl(preProcOptions = list(thresh = 0.8)),
            preProcess = "pca",
            method = "glm"
            )

pred2 <- predict(mod2, testing_subset)

result2 <- confusionMatrix(pred2, testing_subset$diagnosis)

result1
result2
```

## A5

I can't make it work. this exercise will be lacking. :/
