---
title: "Quiz 03"
author: "Matheus Cardoso"
date: "Jun 12, 2020"
output: 
  html_document: 
    fig_caption: yes
    highlight: zenburn
    keep_md: yes
    theme: simplex
    toc: yes
    toc_float: yes
  pdf_document: 
    toc: yes
    fig_caption: yes
    highlight: zenburn
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE, cache = TRUE, warning = FALSE, message = FALSE)

library(caret)
library(tidyverse)
library(rattle)
```

# Q1 

For this quiz we will be using several R packages. R package versions change over time, the right answers have been checked using the following versions of the packages.

AppliedPredictiveModeling: v1.1.6

caret: v6.0.47

ElemStatLearn: v2012.04-0

pgmm: v1.1

rpart: v4.1.8

If you aren't using these versions of the packages, your answers may not exactly match the right answer, but hopefully should be close.

Load the cell segmentation data from the AppliedPredictiveModeling package using the commands: 

```{r q1}
library(AppliedPredictiveModeling)
data(segmentationOriginal)
library(caret)
```

1. Subset the data to a training set and testing set based on the Case variable in the data set.

2. Set the seed to 125 and fit a CART model with the rpart method using all predictor variables and default caret settings.

3. In the final model what would be the final model prediction for cases with the following variable values:

    a. TotalIntench2 = 23,000; FiberWidthCh1 = 10; PerimStatusCh1=2
    
    b. TotalIntench2 = 50,000; FiberWidthCh1 = 10;VarIntenCh4 = 100
    
    c. TotalIntench2 = 57,000; FiberWidthCh1 = 8;VarIntenCh4 = 100
    
    d. FiberWidthCh1 = 8;VarIntenCh4 = 100; PerimStatusCh1=2 

## A1

```{r a1}
set.seed(125)

names(segmentationOriginal)
dim(segmentationOriginal)

inTrain <- createDataPartition(segmentationOriginal$Case, p = .7, list = F)
training <- segmentationOriginal[ inTrain, ]
testing  <- segmentationOriginal[-inTrain, ]

a1_model <- train(Case ~ ., data = training, method = "rpart")
a1_model$finalModel
fancyRpartPlot(a1_model$finalModel)

```

# Q2

If K is small in a K-fold cross validation is the bias in the estimate of out-of-sample (test set) accuracy smaller or bigger? If K is small is the variance in the estimate of out-of-sample (test set) accuracy smaller or bigger. Is K large or small in leave one out cross validation? 

## A2

The bias is larger and the variance is smaller. Under leave one out cross validation K is equal to the sample size. 

# Q3

Load the olive oil data using the commands:

```{r q3}
library(pgmm)
data(olive)
olive = olive[,-1]
```

(NOTE: If you have trouble installing the pgmm package, you can download the -code-olive-/code- dataset here: olive_data.zip. After unzipping the archive, you can load the file using the -code-load()-/code- function in R.)

These data contain information on 572 different Italian olive oils from multiple regions in Italy. Fit a classification tree where Area is the outcome variable. Then predict the value of area for the following data frame using the tree command with all defaults

```{r a3}
newdata = as.data.frame(t(colMeans(olive)))

str(newdata)

a2_model <- train(Area ~ ., data = olive, method = "rpart")
predict(a2_model)
```

## A3

2.783. It is strange because Area should be a qualitative variable - but tree is reporting the average value of Area as a numeric variable in the leaf predicted for newdata

# Q4

Load the South Africa Heart Disease Data and create training and test sets with the following code:

```{r q4}
library(ElemStatLearn)
data(SAheart)
set.seed(8484)
train = sample(1:dim(SAheart)[1],size=dim(SAheart)[1]/2,replace=F)
trainSA = SAheart[train,]
testSA = SAheart[-train,]
```

Then set the seed to 13234 and fit a logistic regression model (method="glm", be sure to specify family="binomial") with Coronary Heart Disease (chd) as the outcome and age at onset, current alcohol consumption, obesity levels, cumulative tabacco, type-A behavior, and low density lipoprotein cholesterol as predictors. Calculate the misclassification rate for your model using this function and a prediction on the "response" scale:
```{r}
missClass = function(values,prediction){sum(((prediction > 0.5)*1) != values)/length(values)}
```

```{r a4}
set.seed(13234)
a4_model <- train(
                chd ~ age + alcohol + obesity + tobacco + typea + ldl,
                data = trainSA,
                method = "glm",
                family = "binomial"
)

missClass(trainSA, predict(a4_model, newdata = trainSA))
missClass(testSA, predict(a4_model, newdata = testSA))
```

## A4

This code does not seems to work, even it's apparently right. maybe the data changed because I needed to install the `ElemStatLearn` package from github? (this package isn't available from CRAN anymore...)

# Q5

Load the Alzheimer's disease data using the commands:

```{r q5}
library(caret)
library(AppliedPredictiveModeling)
set.seed(3433)
data(AlzheimerDisease)
adData = data.frame(diagnosis,predictors)
inTrain = createDataPartition(adData$diagnosis, p = 3/4)[[1]]
training = adData[ inTrain,]
testing = adData[-inTrain,]
```

Create a training data set consisting of only the predictors with variable names beginning with IL and the diagnosis. Build two predictive models, one using the predictors as they are and one using PCA with principal components explaining 80% of the variance in the predictors. Use method="glm" in the train function.

What is the accuracy of each method in the test set? Which is more accurate?

```{r a5}
str(training)
names(training)

training_subset <- training %>% 
                        select(starts_with("IL") | "diagnosis")

testing_subset <- testing %>% 
                        select(starts_with("IL") | "diagnosis")

names(training_subset)
names(testing_subset)

# model Non-PCA
mod1 <- train(
            diagnosis ~ .,
            data = training_subset,
            method = "glm"
            )

pred1 <- predict(mod1, testing_subset)

result1 <- confusionMatrix(pred1, testing_subset$diagnosis)

# model PCA
mod2 <- train(
            diagnosis ~ .,
            data = training_subset,
            trControl = trainControl(preProcOptions = list(thresh = 0.8)),
            preProcess = "pca",
            method = "glm"
            )

pred2 <- predict(mod2, testing_subset)

result2 <- confusionMatrix(pred2, testing_subset$diagnosis)

result1
result2
```

## A5

I can't make it work. this exercise will be lacking. :/
