---
title: "Simulating Exponential Distributions"
subtitle: "A comparison between exponential distributions and the Central Limit Theorem"
author: "Matheus Cardoso"
date: "Jun 04, 2020"
output: 
  html_document: 
    fig_caption: yes
    highlight: zenburn
    keep_md: yes
    theme: simplex
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE, cache = FALSE, warning = FALSE, message = FALSE)
```

# Overview

In this project I aim to present you how a randomly generated exponential distribution sample relates to the Central Limit Theorem (CLT from now on).
I'll present basicaly 3 points related to this theme:

 1. Show the sample mean and compare it to the theoretical mean of the distribution.
 2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
 3. Show that the distribution is approximately normal.
 
## Setting up the enviroment

Before we start, we need to setup the enviroment

```{r load_libs}
library(tidyverse)
```

Now, we can start to analyse the distribuitions.

# Simulations

## Sample Mean versus Theoretical Mean

## Sample Variance versus Theoretical Variance

## Distribution
