---
title: "Data Analysis of Tooth Growth"
subtitle: "The Effect of Vitamin C on Tooth Growth in Guinea Pigs"
author: "Matheus Cardoso"
date: "Jun 04, 2020"
output: 
  html_document: 
    fig_caption: yes
    highlight: zenburn
    keep_md: yes
    theme: simplex
    toc: yes
    toc_float: yes
  pdf_document: 
    toc: yes
    fig_caption: yes
    highlight: zenburn
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE, cache = FALSE, warning = FALSE, message = FALSE)
```

# Abstract

In this report we will explore some basic informations about tooth growth related to different doses of vitamin C in pigs.
I'll briefly describe some exploratory data analysis (EDA) performed, and after that present the results and a short discussion.

# Methods

We start setting the environment and loading the ToothGrowth dataset, a built-in dataframe from R.
This is the description of the data:

> <sub>The response is the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, orange juice or ascorbic acid (a form of vitamin C and coded as VC).</sub>


## Exploratory Data Analysis

```{r setenv}
library(tidyverse)
library(pairwiseComparisons)

# for reproducibility we need to set seed
set.seed(42)
```

```{r load_data}
df <- as_tibble(ToothGrowth) %>%
    mutate(supp = case_when(
        supp == "VC" ~ "Vitamin C",
        supp == "OJ" ~ "Orange Juice"
    )) %>% 
    group_by(supp, dose)

knitr::kable(head(sample_n(df, 1)))
```


## Visualy Analysing Data

```{r data_plot1}
ggplot(df) +
    geom_boxplot(aes(supp, len, fill = supp)) +
    facet_wrap( ~ dose) +
    labs(
        title = "Comparison of Tooth Growth by Delivery Method",
        subtitle = "Note how dose levels affect length",
        x = "Delivery Method",
        y = "Tooth Length",
        caption = "Note")
```

## Hypothesis

As we can see in the plots there seams to exist a relationship in regard to the dose levels.
It's harder to say the same about the two different delivery methods.
Let's perform some statistical analysis to assess the significance for this two types of variance:

 - Assess if dose levels are significant to tooth growth.
 - Assess if delivery methods are significant to tooth growth.


```{r data_comparisons}
hypothesis_dose_level <- t.test(df$len ~ df$supp, paired = FALSE)
hypothesis_dose_level
```


## Conclusion

Reading the result of this analysis, we can conclude that it's indeed statistically significant the _dose_ variable,
i.e. as we increase the dose, so does the length of the teeth. 
In the other hand, we can't say the same about how vitamin C is delivered.  
