---
title: "Data Analysis of Cars Performance | Motor Trend Magazine"
subtitle: "How does transmission type inffluences Miles Per Gallon?"
author: "Matheus Cardoso"
date: "Jun 10, 2020"
output: 
  html_document: 
    fig_caption: yes
    highlight: zenburn
    keep_md: yes
    theme: simplex
    toc: yes
    code_folding: hide
    toc_float: yes
  pdf_document: 
    toc: yes
    fig_caption: yes
    highlight: zenburn
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE, cache = FALSE, warning = FALSE, message = FALSE)
```

# Overview

In this post we'll perform some data analysis in regard to transmission type and miles per gallon (MPG).
We'll seek to answer 2 main questions:

 - “Is an automatic or manual transmission better for MPG”? 
 
 - How can we "Quantify the MPG difference between automatic and manual transmissions"?

# Data Exploration

## Introduction to dataset used

We are going to use a famous dataset in R, called `mtcars`.
Here is a brief explanation about it:

><sub>The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).</sub>

><sub>Format:
A data frame with 32 observations on 11 (numeric) variables.
<br />[, 1]	mpg	Miles/(US) gallon
<br />[, 2]	cyl	Number of cylinders
<br />[, 3]	disp	Displacement (cu.in.)
<br />[, 4]	hp	Gross horsepower
<br />[, 5]	drat	Rear axle ratio
<br />[, 6]	wt	Weight (1000 lbs)
<br />[, 7]	qsec	1/4 mile time
<br />[, 8]	vs	Engine (0 = V-shaped, 1 = straight)
<br />[, 9]	am	Transmission (0 = automatic, 1 = manual)
<br />[,10]	gear	Number of forward gears
<br />[,11]	carb	Number of carburetors</sub>


## Load data and process it

In this section we load libraries, load data and modify it to a more tidy form.

```{r load_data_and_libs}
library(tidyverse)
library(magrittr)
library(GGally)
library(knitr)
library(glue)

# set seed for reproducibility
set.seed(1)

d <- mtcars %>%
        as_tibble() %>% 
        mutate(am = factor(am, labels = c("Automatic", "Manual")),
               vs = factor(vs, labels = c("V", "S")),
               cyl = factor(cyl))

summary_d <- tibble(
    "Number of Rows" = nrow(d),
    "Number of Colunms" = ncol(d)
)

# show dataframe info
kable(
    summary_d,
    caption = "A summary of the dimensions of `mtcars` dataset"
    )

kable(
    sample_n(d, 5),
    caption = "A quick look at the raw data"
)
```


## Basic visualization of the data

Here we show some basic plots demonstrating properties of the data.


```{r plots}

ggplot(d) +
    geom_point(aes(mpg, disp, color = am)) +
    geom_smooth(aes(mpg, disp, color = am), method = "lm") +
    labs(
        title = "Variation of Engine Displacement as Miles per Galon as increases",
        caption = "Note how transmission type influences the relationship",
        x = "Miles per Galon",
        y = "Engine Displacement (cu.in.)"
    ) +
    scale_color_discrete(name = "Transmission Type")

ggplot(d) +
    geom_point(aes(mpg, gear, color = am)) +
    geom_smooth(aes(mpg, gear, color = am), method = "lm") +
    labs(
        title = "How Gear type and Miles per Galon are related",
        caption = "Note how automatic cars tend to have a positive association between Miles per Galon and Gear.
        The same does not occour with manual cars.",
        x = "Miles per Galon",
        y = "Gear type"
    ) +
    scale_color_discrete(name = "Transmission Type")


plot_t.test <- t.test(d$mpg ~ d$am)
ggplot(d) +
    geom_boxplot(aes(am, mpg, fill = am)) +
    scale_color_discrete(name = "Transmission Type") +
    labs(
        title = "How does Transmission Type influences Miles per Galon?",
        x = "Transmission Type",
        y = "Miles per Galon",
        caption = glue(
            "As you can see, the difference between transmission type is quite large.
            The p-value for this two groups is {round(plot_t.test$p.value, 3)}
            "
        )
    )
```

```{r model_data}

foo <- lm(mpg ~ ., data = d)

# a <- d %>% select(mpg, disp, hp, drat, wt, qsec, am)
# ggpairs(a, mapping = ggplot2::aes(color = am))
# summary(foo)
# ggplot(foo)
```
